---
title: "final_project"
author: "Annie Clark"
date: "December 8, 2018"
output: github_document
---

1) Lit review
2) EDA/tidying
3) lit review and eda to help with variable selection


## missing values

```{r}
library(tidyverse)

cancer_data = read_csv(file = "./data/Cancer_Registry.csv") %>% 
  janitor::clean_names()

names(cancer_data)

cancer_data %>% 
  filter(is.na(avg_ann_count)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(avg_deaths_per_year)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(target_death_rate)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(incidence_rate)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(med_income)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pop_est2015)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(poverty_percent)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(study_per_cap)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(binned_inc)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(median_age)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(median_age_male)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(median_age_female)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(geography)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(avg_household_size)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(percent_married)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_no_hs18_24)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_hs18_24)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_some_col18_24)) %>% 
  nrow()

##2285

cancer_data %>% 
  filter(is.na(pct_bach_deg18_24)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_hs25_over)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_private_coverage)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_employed16_over)) %>% 
  nrow()

## 152

cancer_data %>% 
  filter(is.na(pct_unemployed16_over)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_private_coverage)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_private_coverage_alone)) %>% 
  nrow()

##609

cancer_data %>% 
  filter(is.na(pct_emp_priv_coverage)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_public_coverage)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_public_coverage_alone)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_white)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_black)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_asian)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_other_race)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(pct_married_households)) %>% 
  nrow()

cancer_data %>% 
  filter(is.na(birth_rate)) %>% 
  nrow()

```

variables with missing data:
- pct_private_coverage_alone (609 missing) 20%
- pct_employed16_over (152 missing) 5%
- pct_some_col18_24 (2285 missing) 75%


## correlation

```{r}


summary(cancer_data)

## binned_inc - character
## geography - character

cancer_data %>% 
  select(-binned_inc, -geography) %>% ##remove characters
  cor() %>% 
  as_tibble()

library(corrplot)

cancer_data %>% 
  select(-binned_inc, -geography,
         -pct_private_coverage_alone, -pct_employed16_over, -pct_some_col18_24) %>% ##remove characters
  cor() %>% 
  corrplot(., method = "square", insig = "blank", diag=FALSE, number.cex = .7)

```




There are clusters of variables that look to be correlated with other variables. 

If we choose to use any other following variables, we need to check what other variables are in the model to make sure there is not a multicollinearity issue:

* med_income
* poverty_percent
* pct_public_coverage_along
* pct_public_coverage
* pct_emp_priv_coverage
* pct_private_coverage

Also, noteable pairs highly correlated:

* pct_black & pct_white
* pct_married_household & avg_household_size
... there are a dozen more


## scatterplots of target_death_rate vs. risk factors 
```{r}
cancer_data %>% 
  ggplot(aes(x = avg_ann_count, y = target_death_rate)) + 
  geom_point()

## looks like needs a log transformation

cancer_data %>% 
  ggplot(aes(x = log(avg_ann_count), y = target_death_rate)) + 
  geom_point()

## looks curvilinear


cancer_data %>% 
  ggplot(aes(x = avg_deaths_per_year, y = target_death_rate)) + 
  geom_point()

## looks like needs a log transformation


cancer_data %>% 
  ggplot(aes(x = log(avg_deaths_per_year), y = target_death_rate)) + 
  geom_point()

## still looks curvilinear


cancer_data %>% 
  ggplot(aes(x = incidence_rate, y = target_death_rate)) + 
  geom_point()

## linear, with some x extreme outliers

cancer_data %>% 
  ggplot(aes(x = med_income, y = target_death_rate)) + 
  geom_point()

## looks a little curvilinear -- need to do transformation

cancer_data %>% 
  ggplot(aes(x = (1/med_income), y = target_death_rate)) + 
  geom_point()

## looks linear...

cancer_data %>% 
  ggplot(aes(x = pop_est2015, y = target_death_rate)) + 
  geom_point()

## need to do log transformation

cancer_data %>% 
  ggplot(aes(x = log(pop_est2015), y = target_death_rate)) + 
  geom_point()

## still looks a little curvilinear

cancer_data %>% 
  ggplot(aes(x = poverty_percent, y = target_death_rate)) + 
  geom_point()

## looks linear

cancer_data %>% 
  ggplot(aes(x = study_per_cap, y = target_death_rate)) + 
  geom_point()

## need to do a log transformation

cancer_data %>% 
  ggplot(aes(x = log(study_per_cap), y = target_death_rate)) + 
  geom_point()

## looks linear, but also has a weird 0 varlies all along the y values

cancer_data %>% 
  ggplot(aes(x = binned_inc, y = target_death_rate)) + 
  geom_point()

## looks randomly scattered throughout the bins.. should do boxplot

cancer_data %>% 
  ggplot(aes(x = binned_inc,  y = target_death_rate)) + 
  geom_boxplot() + 
  theme(axis.text.x=element_text(angle=45,hjust=1)) 

cancer_data %>% 
  group_by(binned_inc) %>% 
  summarise(n = n())

## oddly a similar number of observations per bin... 

cancer_data %>% 
  ggplot(aes(x = median_age, y = target_death_rate)) + 
  geom_point()

## this looks like an issue - let's look further into median_age

cancer_data %>% 
  group_by(round(median_age)) %>% 
  summarize(n = n()) 

## there are ages that go from 22 - 65 and then from 349 - 624. This is an error with the data

cancer_data %>% 
  ggplot(aes(x = median_age_male, y = target_death_rate)) + 
  geom_point()

## looks like curvilinear, let's try transformation

cancer_data %>% 
  ggplot(aes(x = median_age_male^2, y = target_death_rate)) + 
  geom_point()

## tried ^2, ^3, inverse, log, nothing really makes it look better

cancer_data %>% 
  ggplot(aes(x = median_age_female, y = target_death_rate)) + 
  geom_point()

## similar to male median age


cancer_data %>% 
  ggplot(aes(x = median_age_female^2, y = target_death_rate)) + 
  geom_point()

## tried ^2, ^3, inverse, log, nothing really makes it look better



cancer_data %>% 
  ggplot(aes(x = avg_household_size, y = target_death_rate)) + 
  geom_point()

## there are a ton of 0 values; this coule potentially be a data quality issue

cancer_data %>% 
  count(avg_household_size)

## it looks like there are a ton of 0.02.... not sure what this means?


cancer_data %>% 
  ggplot(aes(x = percent_married, y = target_death_rate)) + 
  geom_point()

## looks linear; potentially curilinear


cancer_data %>% 
  ggplot(aes(x = percent_married^2, y = target_death_rate)) + 
  geom_point()

## looks better

cancer_data %>% 
  ggplot(aes(x = pct_no_hs18_24, y = target_death_rate)) + 
  geom_point()

## looks linear; potentially curilinear


cancer_data %>% 
  ggplot(aes(x = pct_hs18_24, y = target_death_rate)) + 
  geom_point()

cancer_data %>% 
  ggplot(aes(x = pct_some_col18_24, y = target_death_rate)) + 
  geom_point()

## linear, but missing values

cancer_data %>% 
  ggplot(aes(x = pct_bach_deg18_24, y = target_death_rate)) + 
  geom_point()

## log transformation

cancer_data %>% 
  ggplot(aes(x = log(pct_bach_deg18_24), y = target_death_rate)) + 
  geom_point()

## makes it a little better, but zero values are an issue (inverse and squared do not work)


cancer_data %>% 
  ggplot(aes(x = pct_hs25_over, y = target_death_rate)) + 
  geom_point()

## looks linear

cancer_data %>% 
  ggplot(aes(x = pct_bach_deg25_over, y = target_death_rate)) + 
  geom_point()

## looks a little curvilinear

cancer_data %>% 
  ggplot(aes(x = log(pct_bach_deg25_over), y = target_death_rate)) + 
  geom_point()

## looks good

cancer_data %>% 
  ggplot(aes(x = pct_employed16_over, y = target_death_rate)) + 
  geom_point()

## looks linear

cancer_data %>% 
  ggplot(aes(x = pct_unemployed16_over, y = target_death_rate)) + 
  geom_point()

## looks linear

cancer_data %>% 
  ggplot(aes(x = pct_private_coverage, y = target_death_rate)) + 
  geom_point()

## looks linear

cancer_data %>% 
  ggplot(aes(x = pct_private_coverage_alone, y = target_death_rate)) + 
  geom_point()

## looks linear

cancer_data %>% 
  ggplot(aes(x = pct_emp_priv_coverage, y = target_death_rate)) + 
  geom_point()

## looks linear but also heteroscedastic

cancer_data %>% 
  ggplot(aes(x = pct_public_coverage, y = target_death_rate)) + 
  geom_point()

## looks linear

cancer_data %>% 
  ggplot(aes(x = pct_public_coverage_alone, y = target_death_rate)) + 
  geom_point()

## looks linear

cancer_data %>% 
  ggplot(aes(x = pct_white, y = target_death_rate)) + 
  geom_point()

## log transformation needed


cancer_data %>% 
  ggplot(aes(x = 1/pct_white, y = target_death_rate)) + 
  geom_point()

## does not look linear, log, inverse and quadratics don't work

cancer_data %>% 
  ggplot(aes(x = pct_black, y = target_death_rate)) + 
  geom_point()

## need tranformation


cancer_data %>% 
  ggplot(aes(x = log(pct_black), y = target_death_rate)) + 
  geom_point()

## log looks better, but negative values may be an issue


cancer_data %>% 
  ggplot(aes(x = pct_asian, y = target_death_rate)) + 
  geom_point()

## need transformation

cancer_data %>% 
  ggplot(aes(x = log(pct_asian), y = target_death_rate)) + 
  geom_point()

## log looks better, but negative values may be an issue

cancer_data %>% 
  ggplot(aes(x = pct_other_race, y = target_death_rate)) + 
  geom_point()

## need transformation

cancer_data %>% 
  ggplot(aes(x = log(pct_other_race), y = target_death_rate)) + 
  geom_point()

## log looks better, but negative values may be an issue

cancer_data %>% 
  ggplot(aes(x = pct_married_households, y = target_death_rate)) + 
  geom_point()

## looks linear

cancer_data %>% 
  ggplot(aes(x = birth_rate, y = target_death_rate)) + 
  geom_point()

## data points are a cluster


cancer_data %>% 
  ggplot(aes(x = (birth_rate^3), y = target_death_rate)) + 
  geom_point()

## transformations don't really work

```


## Non linear variables

* avg_ann_count (transformations don't work)
* avg_deaths_per_year (transformations don't work)
* med_income (inverse transformation worked)
* study_per_cap (log transformation worked, but still has weird 0 values)
* median_age_male (transformations don't work)
* percent_married (squared works)
* pct_no_hs18_24 (potentially curvlinear, squared/log do not work)
* pct_bach_deg18_24 (log makes a little better, but zeros are an issue)
* pct_bach_deg25_over (log works)
* pct_white (transformation don't work)
* pct_black (log works, but negative values clustered)
* pct_asian (log works, but negative values clustered)
* birth_rate (all values are clustered)


## Variables with data quality concerns

* median_age (median ages of 300 - 600)


## Other things to note:

* incidence_rate has some extreme x outliers
* avg_household_size lots of median households of 0.02 size... not sure what this means



forward selection
backward selection
stepwise (forward/backward)
criterion-based
ridge
lasso


```{r}
library(tidyverse)

cancer_data_state =
  cancer_data %>% 
  separate(geography, into = c("county", "state"), sep = ",")

cancer_data_state = 
  cancer_data_state %>% 
  select (-median_age, -binned_inc, -pct_some_col18_24, 
          -pct_private_coverage_alone, pct_employed16_over)

cancer_data_clean = 
  cancer_data %>% 
  separate(geography, into = c("county", "state"), sep = ",") %>% 
  select (-median_age, -binned_inc, -pct_some_col18_24, 
          -pct_private_coverage_alone, pct_employed16_over,
          -county)


step_fit <- lm(target_death_rate ~ ., data = cancer_data_clean)

summary(step_fit)

step(step_fit, direction='backward')
step(step_fit, direction='forward')


cancer_data_grouped = 
  cancer_data_state %>% 
  group_by(state) %>% 
  summarise(avg_ann_count = mean(avg_ann_count),
            incidence_rate = mean(incidence_rate),
            poverty_percent = mean(poverty_percent),
            pct_no_hs18_24 = mean(pct_no_hs18_24),
            pct_bach_deg18_24 = mean(pct_bach_deg18_24),
            pct_employed16_over = mean(pct_employed16_over),
            pct_public_coverage_alone = mean(pct_public_coverage_alone),
            pct_asian = mean(pct_asian),
            birth_rate = mean(birth_rate),
            avg_deaths_per_year = mean(avg_deaths_per_year),
            med_income = mean(med_income),
            study_per_cap = mean(study_per_cap),
            median_age_male = mean(median_age_male),
            avg_household_size = mean(avg_household_size),
            pct_hs18_24 = mean(pct_hs18_24),
            pct_hs25_over = mean(pct_hs25_over),
            pct_unemployed16_over = mean(pct_unemployed16_over),
            pct_emp_priv_coverage = mean(pct_emp_priv_coverage),
            pct_white = mean(pct_white),
            pct_other_race = mean(pct_other_race),
            target_death_rate = mean(target_death_rate),
            pop_est2015 = mean(pop_est2015),
            median_age_female = mean(median_age_female),
            percent_married = mean(percent_married),
            pct_bach_deg25_over = mean(pct_bach_deg25_over),
            pct_private_coverage = mean(pct_private_coverage),
            pct_public_coverage = mean(pct_public_coverage),
            pct_black = mean(pct_black),
            pct_married_households = mean(pct_married_households))


names(cancer_data)




step_fit <- lm(target_death_rate ~ ., data = cancer_data_grouped)

summary(step_fit)

cor(cancer_data_grouped)

step(step_fit, direction='backward')
step(step_fit, direction='forward')
```
- pct_private_coverage_alone (609 missing) 20%
- pct_employed16_over (152 missing) 5%
- pct_some_col18_24 



```{r}

cancer_data_grouped %>% 
  select(-state) %>% ##remove characters
  cor() %>% 
  as_tibble()

library(corrplot)

cancer_data_grouped %>% 
  select(-state,
         -pct_employed16_over) %>% ##remove characters
  cor() %>% 
  corrplot(., method = "square", insig = "blank", diag=FALSE, number.cex = .7)


step_fit <- lm(target_death_rate ~ ., groups = state, data = cancer_data_grouped)
summary(step_fit)


step(step_fit, direction='backward')
step(step_fit, direction='forward')

```


```{r}
step_fit <- lm(target_death_rate ~ avg_ann_count, avg_deaths_per_year , 
               groups = state, data = cancer_data_clean)

names(cancer_data_clean)


summary(step_fit)

cor(cancer_data_grouped)

step(step_fit, direction='backward')
step(step_fit, direction='forward')


install.packages("glmnet", repos = "http://cran.us.r-project.org")
library(glmnet)
test.mat = model.matrix(target_death_rate ~., data = cancer_data_clean)[, -1]
lasso.mod = glmnet(test.mat, target_death_rate, alpha = 1)

set.seed(1)

cv.out = cv.glmnet(test.mat,target_death_rate ,alpha=1)
plot(cv.out)

library(leaps)

Forward Stepwise Selection
regfit.full = regsubsets(target_death_rate ~ ., data=cancer_data_clean, method ="forward")
summary(regfit.full)
reg.summary=summary(regfit.full)

plot(reg.summary$rss ,xlab="Number of Variables ",ylab="RSS",
type="l")

plot(reg.summary$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")





```

