---
title: "annie_models"
author: "Annie Clark"
date: "December 15, 2018"
output: github_document
---


```{r}
library(tidyverse)
library(HH)
library(leaps)
library(modelr)

```


```{r}
library(tidyverse)

cancer_data = read_csv(file = "./data/Cancer_Registry.csv") %>% 
  janitor::clean_names()

cancer_data_grouped = 
  cancer_data %>% 
  separate(geography, into = c("county", "state"), sep = ",") %>% 
  group_by(state) %>% 
  select (-median_age, -binned_inc, -pct_some_col18_24, 
          -pct_private_coverage_alone, pct_employed16_over, -county) %>% 
  summarise(avg_ann_count = mean(avg_ann_count),
            incidence_rate = mean(incidence_rate),
            poverty_percent = mean(poverty_percent),
            pct_no_hs18_24 = mean(pct_no_hs18_24),
            pct_bach_deg18_24 = mean(pct_bach_deg18_24),
            ##pct_employed16_over = mean(pct_employed16_over),
            pct_public_coverage_alone = mean(pct_public_coverage_alone),
            pct_asian = mean(pct_asian),
            birth_rate = mean(birth_rate),
            avg_deaths_per_year = mean(avg_deaths_per_year),
            med_income = mean(med_income),
            study_per_cap = mean(study_per_cap),
            median_age_male = mean(median_age_male),
            avg_household_size = mean(avg_household_size),
            pct_hs18_24 = mean(pct_hs18_24),
            pct_hs25_over = mean(pct_hs25_over),
            pct_unemployed16_over = mean(pct_unemployed16_over),
            pct_emp_priv_coverage = mean(pct_emp_priv_coverage),
            pct_white = mean(pct_white),
            pct_other_race = mean(pct_other_race),
            target_death_rate = mean(target_death_rate),
            pop_est2015 = mean(pop_est2015),
            median_age_female = mean(median_age_female),
            percent_married = mean(percent_married),
            pct_bach_deg25_over = mean(pct_bach_deg25_over),
            pct_private_coverage = mean(pct_private_coverage),
            pct_public_coverage = mean(pct_public_coverage),
            pct_black = mean(pct_black),
            pct_married_households = mean(pct_married_households)) %>% 
  select(-state)

## median for everyone; 

library(leaps)

## best subset selection

regfit.full = regsubsets(target_death_rate ~ ., data= cancer_data_grouped, nvmax = 27)
reg.summary = summary(regfit.full)
##reg.summary

par(mfrow=c(2, 2))
plot(reg.summary$rss ,xlab="Number of Variables ",ylab="RSS", type="l")

plot(reg.summary$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")

plot(reg.summary$cp,xlab="Number of Variables ",ylab="Cp",type="l")

plot(reg.summary$bic ,xlab="Number of Variables ",ylab="BIC",type="l")

which.min(reg.summary$rss)
which.max(reg.summary$adjr2)
which.min(reg.summary$cp)
which.min(reg.summary$bic)

## forward step wise 

regfit.forward = regsubsets(target_death_rate ~ ., data = cancer_data_grouped, 
                            method ="forward", nvmax = 27)
reg.summary=summary(regfit.forward)
##reg.summary

par(mfrow=c(2, 2))
plot(reg.summary$rss ,xlab="Number of Variables ",ylab="RSS", type="l")

plot(reg.summary$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")

plot(reg.summary$cp,xlab="Number of Variables ",ylab="Cp",type="l")

plot(reg.summary$bic ,xlab="Number of Variables ",ylab="BIC",type="l")

which.min(reg.summary$rss)
which.max(reg.summary$adjr2)
which.min(reg.summary$cp)
which.min(reg.summary$bic)


## backward step wise 

regfit.backward = regsubsets(target_death_rate ~ ., data= cancer_data_grouped, 
                             method = "backward", nvmax = 27)
reg.summary = summary(regfit.backward)
##reg.summary

par(mfrow=c(2, 2))
plot(reg.summary$rss ,xlab="Number of Variables ",ylab="RSS", type="l")

plot(reg.summary$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")

plot(reg.summary$cp,xlab="Number of Variables ",ylab="Cp",type="l")

plot(reg.summary$bic ,xlab="Number of Variables ",ylab="BIC",type="l")

which.min(reg.summary$rss)
which.max(reg.summary$adjr2)
which.min(reg.summary$cp)
which.min(reg.summary$bic)



## lasso 

require(methods)
attach(cancer_data_grouped)

library(glmnet)
test.mat = model.matrix(target_death_rate ~ ., data = cancer_data_grouped)[, -1]
lasso.mod = glmnet(test.mat, target_death_rate, alpha = 1)
set.seed(1)
cv.out = cv.glmnet(test.mat,target_death_rate,alpha=1)
par(mfrow=c(1, 1))
plot(cv.out)

bestlam = cv.out$lambda.min
bestlam

out = glmnet (test.mat, target_death_rate, alpha=1)
lasso.coef = predict (out, type="coefficients", s= bestlam)
lasso.coef

rsq = 1 - cv.out$cvm/var(target_death_rate)

summary(lasso.mod)

lasso.mod



## ridge regression

test.mat = model.matrix(target_death_rate ~ ., data = cancer_data_grouped)[, -1]
ridge.mod=cv.glmnet(test.mat, target_death_rate, alpha=0, thresh = 1e-12)
bestlam =ridge.mod$lambda.min
par(mfrow=c(1, 1))
plot(ridge.mod)
ridge.mod$lambda.min

coef(ridge.mod)



## split into test/training datasets

library(ISLR)
train = sample(1:nrow(cancer_data_grouped), nrow(cancer_data_grouped)/2)
test = (-train)
test.data = cancer_data_grouped[test,]
train.data = cancer_data_grouped[train,]


```


```{r}
library(tidyverse)

cancer_data = read_csv(file = "./data/Cancer_Registry.csv") %>% 
  janitor::clean_names()

cancer_data_clean = 
  cancer_data %>% 
  separate(geography, into = c("county", "state"), sep = ",") %>% 
  select (-median_age, -binned_inc, -pct_some_col18_24, 
          -pct_private_coverage_alone, pct_employed16_over, -county, -state) 


library(leaps)

## best subset selection

regfit.full = regsubsets(target_death_rate ~ ., data= cancer_data_clean, nvmax = 27)
reg.summary = summary(regfit.full)
##reg.summary

par(mfrow=c(2, 2))
plot(reg.summary$rss ,xlab="Number of Variables ",ylab="RSS", type="l")

plot(reg.summary$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")

plot(reg.summary$cp,xlab="Number of Variables ",ylab="Cp",type="l")

plot(reg.summary$bic ,xlab="Number of Variables ",ylab="BIC",type="l")

which.min(reg.summary$rss)
which.max(reg.summary$adjr2)
which.min(reg.summary$cp)
which.min(reg.summary$bic)

## forward step wise 

regfit.forward = regsubsets(target_death_rate ~ ., data = cancer_data_clean, 
                            method ="forward", nvmax = 27)
reg.summary=summary(regfit.forward)
##reg.summary

par(mfrow=c(2, 2))
plot(reg.summary$rss ,xlab="Number of Variables ",ylab="RSS", type="l")

plot(reg.summary$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")

plot(reg.summary$cp,xlab="Number of Variables ",ylab="Cp",type="l")

plot(reg.summary$bic ,xlab="Number of Variables ",ylab="BIC",type="l")

which.min(reg.summary$rss)
which.max(reg.summary$adjr2)
which.min(reg.summary$cp)
which.min(reg.summary$bic)


## backward step wise 

regfit.backward = regsubsets(target_death_rate ~ ., data= cancer_data_clean, 
                             method = "backward", nvmax = 27)
reg.summary = summary(regfit.backward)

##reg.summary

par(mfrow=c(2, 2))
plot(reg.summary$rss ,xlab="Number of Variables ",ylab="RSS", type="l")

plot(reg.summary$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")

plot(reg.summary$cp,xlab="Number of Variables ",ylab="Cp",type="l")

plot(reg.summary$bic ,xlab="Number of Variables ",ylab="BIC",type="l")

which.min(reg.summary$rss)
which.max(reg.summary$adjr2)
which.min(reg.summary$cp)
which.min(reg.summary$bic)

```


```{r}
cancer_data_grouped_median = 
  cancer_data %>% 
  separate(geography, into = c("county", "state"), sep = ",") %>% 
  group_by(state) %>% 
  select (-median_age, -binned_inc, -pct_some_col18_24, 
          -pct_private_coverage_alone, pct_employed16_over, -county) %>% 
  summarise(avg_ann_count = median(avg_ann_count),
            incidence_rate = median(incidence_rate),
            poverty_percent = median(poverty_percent),
            pct_no_hs18_24 = median(pct_no_hs18_24),
            pct_bach_deg18_24 = median(pct_bach_deg18_24),
            pct_public_coverage_alone = median(pct_public_coverage_alone),
            pct_asian = median(pct_asian),
            birth_rate = median(birth_rate),
            avg_deaths_per_year = median(avg_deaths_per_year),
            med_income = median(med_income),
            study_per_cap = median(study_per_cap),
            median_age_male = median(median_age_male),
            avg_household_size = median(avg_household_size),
            pct_hs18_24 = median(pct_hs18_24),
            pct_hs25_over = median(pct_hs25_over),
            pct_unemployed16_over = median(pct_unemployed16_over),
            pct_emp_priv_coverage = median(pct_emp_priv_coverage),
            pct_white = median(pct_white),
            pct_other_race = median(pct_other_race),
            target_death_rate = median(target_death_rate),
            pop_est2015 = median(pop_est2015),
            median_age_female = median(median_age_female),
            percent_married = median(percent_married),
            pct_bach_deg25_over = median(pct_bach_deg25_over),
            pct_private_coverage = median(pct_private_coverage),
            pct_public_coverage = median(pct_public_coverage),
            pct_black = median(pct_black),
            pct_married_households = median(pct_married_households)) %>% 
  select(-state)



regfit.full = regsubsets(target_death_rate ~ ., data= cancer_data_grouped_median, nvmax = 27)
reg.summary = summary(regfit.full)
##reg.summary

par(mfrow=c(2, 2))
plot(reg.summary$rss ,xlab="Number of Variables ",ylab="RSS", type="l")

plot(reg.summary$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")

plot(reg.summary$cp,xlab="Number of Variables ",ylab="Cp",type="l")

plot(reg.summary$bic ,xlab="Number of Variables ",ylab="BIC",type="l")

which.min(reg.summary$rss)
which.max(reg.summary$adjr2)
which.min(reg.summary$cp)
which.min(reg.summary$bic)

## forward step wise 

regfit.forward = regsubsets(target_death_rate ~ ., data = cancer_data_grouped_median, 
                            method ="forward", nvmax = 27)
reg.summary=summary(regfit.forward)
##reg.summary

par(mfrow=c(2, 2))
plot(reg.summary$rss ,xlab="Number of Variables ",ylab="RSS", type="l")

plot(reg.summary$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")

plot(reg.summary$cp,xlab="Number of Variables ",ylab="Cp",type="l")

plot(reg.summary$bic ,xlab="Number of Variables ",ylab="BIC",type="l")

which.min(reg.summary$rss)
which.max(reg.summary$adjr2)
which.min(reg.summary$cp)
which.min(reg.summary$bic)


## backward step wise 

regfit.backward = regsubsets(target_death_rate ~ ., data= cancer_data_grouped, 
                             method = "backward", nvmax = 27)
reg.summary = summary(regfit.backward)
##reg.summary

par(mfrow=c(2, 2))
plot(reg.summary$rss ,xlab="Number of Variables ",ylab="RSS", type="l")

plot(reg.summary$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")

plot(reg.summary$cp,xlab="Number of Variables ",ylab="Cp",type="l")

plot(reg.summary$bic ,xlab="Number of Variables ",ylab="BIC",type="l")

which.min(reg.summary$rss)
which.max(reg.summary$adjr2)
which.min(reg.summary$cp)
which.min(reg.summary$bic)



## lasso 

require(methods)
attach(cancer_data_grouped)

library(glmnet)
test.mat = model.matrix(target_death_rate ~ ., data = cancer_data_grouped)[, -1]
lasso.mod = glmnet(test.mat, target_death_rate, alpha = 1)
set.seed(1)
cv.out = cv.glmnet(test.mat,target_death_rate,alpha=1)
par(mfrow=c(1, 1))
plot(cv.out)

bestlam = cv.out$lambda.min
bestlam

out = glmnet (test.mat, target_death_rate, alpha=1)
lasso.coef = predict (out, type="coefficients", s= bestlam)
lasso.coef

rsq = 1 - cv.out$cvm/var(target_death_rate)

summary(lasso.mod)

lasso.mod



## ridge regression

test.mat = model.matrix(target_death_rate ~ ., data = cancer_data_grouped)[, -1]
ridge.mod=cv.glmnet(test.mat, target_death_rate, alpha=0, thresh = 1e-12)
bestlam =ridge.mod$lambda.min
par(mfrow=c(1, 1))
plot(ridge.mod)
ridge.mod$lambda.min

coef(ridge.mod)

```


```{r}

model_data = 
  cancer_data %>%
  dplyr::select(-pct_some_col18_24) %>%
  na.omit() %>%
  dplyr::select(-c(binned_inc, median_age, avg_deaths_per_year, pop_est2015)) %>% 
  mutate(log_avg_ann_count = log(avg_ann_count),
         inverse_med_income = 1/(med_income),
         sqrt_poverty_percent = sqrt(poverty_percent),
         study_per_cap = study_per_cap + 0.01,
         log_study_per_cap = log(study_per_cap),
         log_avg_household_size = log(avg_household_size),
         sqrt_pct_no_hs18_24 = sqrt(pct_no_hs18_24),
         sqrt_pct_bach_deg18_24 = sqrt(pct_bach_deg18_24),
         log_pct_bach_deg25_over = log(pct_bach_deg25_over),
         sqrt_pct_unemployed16_over = sqrt(pct_unemployed16_over),
         sqrt_pct_public_coverage_alone = sqrt(pct_public_coverage_alone),
         pct_black = pct_black + 0.01,
         log_pct_black = log(pct_black),
         pct_asian = pct_asian + 0.01,
         log_pct_asian = log(pct_asian),
         pct_other_race = pct_other_race + 0.01,
         log_other_race = log(pct_other_race),
         sqrt_birth_rate = sqrt(birth_rate)) %>% 
  dplyr::select(-c(avg_ann_count, med_income, poverty_percent, study_per_cap, 
            avg_household_size, pct_no_hs18_24, pct_bach_deg18_24, pct_bach_deg25_over,
            pct_unemployed16_over, pct_public_coverage_alone, pct_black, 
            pct_asian, pct_other_race, birth_rate))

model_data_state_mean = 
  model_data %>% 
  separate(geography, into = c("county", "state"), sep = ",") %>% 
  group_by(state) %>% 
  dplyr::select (-county) %>% 
  summarise(log_avg_ann_count = mean(log_avg_ann_count),
            incidence_rate = mean(incidence_rate),
            sqrt_poverty_percent = mean(sqrt_poverty_percent),
            sqrt_pct_no_hs18_24 = mean(sqrt_pct_no_hs18_24),
            sqrt_pct_bach_deg18_24 = mean(sqrt_pct_bach_deg18_24),
            sqrt_pct_public_coverage_alone = mean(sqrt_pct_public_coverage_alone),
            log_pct_asian = mean(log_pct_asian),
            sqrt_birth_rate = mean(sqrt_birth_rate),
            inverse_med_income = mean(inverse_med_income),
            log_study_per_cap = mean(log_study_per_cap),
            median_age_male = mean(median_age_male),
            log_avg_household_size = mean(log_avg_household_size),
            pct_hs18_24 = mean(pct_hs18_24),
            pct_hs25_over = mean(pct_hs25_over),
            sqrt_pct_unemployed16_over = mean(sqrt_pct_unemployed16_over),
            pct_emp_priv_coverage = mean(pct_emp_priv_coverage),
            pct_white = mean(pct_white),
            log_other_race = mean(log_other_race),
            target_death_rate = mean(target_death_rate),
            median_age_female = mean(median_age_female),
            percent_married = mean(percent_married),
            log_pct_bach_deg25_over = mean(log_pct_bach_deg25_over),
            pct_private_coverage = mean(pct_private_coverage),
            pct_public_coverage = mean(pct_public_coverage),
            log_pct_black = mean(log_pct_black),
            pct_married_households = mean(pct_married_households), 
            pct_employed16_over = mean(pct_employed16_over),
            pct_private_coverage_alone = mean(pct_private_coverage_alone)) %>% 
  dplyr::select(-state)


model_data_state_median = 
  model_data %>% 
  separate(geography, into = c("county", "state"), sep = ",") %>% 
  group_by(state) %>% 
  dplyr::select (-county) %>% 
  summarise(log_avg_ann_count = median(log_avg_ann_count),
            incidence_rate = median(incidence_rate),
            sqrt_poverty_percent = median(sqrt_poverty_percent),
            sqrt_pct_no_hs18_24 = median(sqrt_pct_no_hs18_24),
            sqrt_pct_bach_deg18_24 = median(sqrt_pct_bach_deg18_24),
            sqrt_pct_public_coverage_alone = median(sqrt_pct_public_coverage_alone),
            log_pct_asian = median(log_pct_asian),
            sqrt_birth_rate = median(sqrt_birth_rate),
            inverse_med_income = median(inverse_med_income),
            log_study_per_cap = median(log_study_per_cap),
            median_age_male = median(median_age_male),
            log_avg_household_size = median(log_avg_household_size),
            pct_hs18_24 = median(pct_hs18_24),
            pct_hs25_over = median(pct_hs25_over),
            sqrt_pct_unemployed16_over = median(sqrt_pct_unemployed16_over),
            pct_emp_priv_coverage = median(pct_emp_priv_coverage),
            pct_white = median(pct_white),
            log_other_race = median(log_other_race),
            target_death_rate = median(target_death_rate),
            median_age_female = median(median_age_female),
            percent_married = median(percent_married),
            log_pct_bach_deg25_over = median(log_pct_bach_deg25_over),
            pct_private_coverage = median(pct_private_coverage),
            pct_public_coverage = median(pct_public_coverage),
            log_pct_black = median(log_pct_black),
            pct_married_households = median(pct_married_households), 
            pct_employed16_over = median(pct_employed16_over),
            pct_private_coverage_alone = median(pct_private_coverage_alone)) %>% 
  dplyr::select(-state)



```


##Mean
```{r}
transform_model = lm(target_death_rate ~ log_avg_ann_count + incidence_rate + 
                    inverse_med_income + sqrt_poverty_percent + median_age_male +
                    median_age_female + log_avg_household_size + percent_married +
                    sqrt_pct_no_hs18_24 + pct_hs18_24 + sqrt_pct_bach_deg18_24 + 
                    pct_hs25_over + log_pct_bach_deg25_over + pct_employed16_over +
                    sqrt_pct_unemployed16_over + pct_private_coverage + 
                    pct_private_coverage_alone + pct_emp_priv_coverage + pct_public_coverage +
                    sqrt_pct_public_coverage_alone + pct_white + pct_married_households +
                    sqrt_birth_rate + log_study_per_cap + log_pct_black + log_pct_asian +
                    log_other_race, data = model_data_state_mean)




vif(transform_model)

correlation_check = model_data %>%
  dplyr::select(inverse_med_income, sqrt_poverty_percent, median_age_male, median_age_female, percent_married, pct_bach_deg25_over, pct_employed16_over, pct_private_coverage, pct_private_coverage_alone, pct_emp_priv_coverage, pct_public_coverage, sqrt_pct_public_coverage_alone, pct_married_households)

library(corrplot)
par(mfrow=c(1, 1))

cor(correlation_check)%>%
  corrplot(method = "square", addCoef.col = "black", tl.col="black", tl.srt=45, insig = "blank", 
         diag=FALSE, number.cex = .7)

## come back and decide which variable to get rid of
```


```{r}
best <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), nvmax=11, ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  


# Select the 'best' model of all subsets for 21-predictor model
round(best(transform_model, nbest = 1), 3)

## plot
regfit.full = regsubsets(target_death_rate ~ log_avg_ann_count + incidence_rate + 
                    inverse_med_income + sqrt_poverty_percent + median_age_male +
                    median_age_female + log_avg_household_size + percent_married +
                    sqrt_pct_no_hs18_24 + pct_hs18_24 + sqrt_pct_bach_deg18_24 + 
                    pct_hs25_over + log_pct_bach_deg25_over + pct_employed16_over +
                    sqrt_pct_unemployed16_over + pct_private_coverage + 
                    pct_private_coverage_alone + pct_emp_priv_coverage + pct_public_coverage +
                    sqrt_pct_public_coverage_alone + pct_white + pct_married_households +
                    sqrt_birth_rate + log_study_per_cap + log_pct_black + log_pct_asian +
                    log_other_race, 
                    data = model_data_state_mean, 
                    nvmax = 10)
reg.summary = summary(regfit.full)

par(mfrow=c(2, 2))
plot(reg.summary$rss ,xlab="Number of Variables ",ylab="RSS", type="l")

plot(reg.summary$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")

plot(reg.summary$cp,xlab="Number of Variables ",ylab="Cp",type="l")

plot(reg.summary$bic ,xlab="Number of Variables ",ylab="BIC",type="l")

## model with 4 variables looks the best
```


##Median
```{r}
transform_model_median = lm(target_death_rate ~ log_avg_ann_count + incidence_rate + 
                    inverse_med_income + sqrt_poverty_percent + median_age_male +
                    median_age_female + log_avg_household_size + percent_married +
                    sqrt_pct_no_hs18_24 + pct_hs18_24 + sqrt_pct_bach_deg18_24 + 
                    pct_hs25_over + log_pct_bach_deg25_over + pct_employed16_over +
                    sqrt_pct_unemployed16_over + pct_private_coverage + 
                    pct_private_coverage_alone + pct_emp_priv_coverage + pct_public_coverage +
                    sqrt_pct_public_coverage_alone + pct_white + pct_married_households +
                    sqrt_birth_rate + log_study_per_cap + log_pct_black + log_pct_asian +
                    log_other_race, data = model_data_state_median)


## come back and decide which variable to get rid of
```

## Best subset selection
```{r}
best <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), nvmax=11, ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  


# Select the 'best' model of all subsets for 21-predictor model
round(best(transform_model_median, nbest = 1), 3)

## plot
regfit_full_mean = regsubsets(target_death_rate ~ log_avg_ann_count + incidence_rate + 
                    inverse_med_income + sqrt_poverty_percent + median_age_male +
                    median_age_female + log_avg_household_size + percent_married +
                    sqrt_pct_no_hs18_24 + pct_hs18_24 + sqrt_pct_bach_deg18_24 + 
                    pct_hs25_over + log_pct_bach_deg25_over + pct_employed16_over +
                    sqrt_pct_unemployed16_over + pct_private_coverage + 
                    pct_private_coverage_alone + pct_emp_priv_coverage + pct_public_coverage +
                    sqrt_pct_public_coverage_alone + pct_white + pct_married_households +
                    sqrt_birth_rate + log_study_per_cap + log_pct_black + log_pct_asian +
                    log_other_race, 
                    data = model_data_state_median, 
                    nvmax = 10)
reg_summary_mean = summary(regfit_full_mean)

par(mfrow=c(2, 2))
plot(reg_summary_median$rss ,xlab="Number of Variables ",ylab="RSS", type="l")

plot(reg_summary_median$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")

plot(reg_summary_median$cp,xlab="Number of Variables ",ylab="Cp",type="l")

plot(reg_summary_median$bic ,xlab="Number of Variables ",ylab="BIC",type="l")
```

Mean is performing better


## Forward subset selection
```{r}
best_forward <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), method = "forward", nvmax=11, ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  


# Select the 'best' model of all subsets for 21-predictor model
round(best_forward(transform_model, nbest = 1), 3)

## plot
regfit_full_mean = regsubsets(target_death_rate ~ log_avg_ann_count + incidence_rate + 
                    inverse_med_income + sqrt_poverty_percent + median_age_male +
                    median_age_female + log_avg_household_size + percent_married +
                    sqrt_pct_no_hs18_24 + pct_hs18_24 + sqrt_pct_bach_deg18_24 + 
                    pct_hs25_over + log_pct_bach_deg25_over + pct_employed16_over +
                    sqrt_pct_unemployed16_over + pct_private_coverage + 
                    pct_private_coverage_alone + pct_emp_priv_coverage + pct_public_coverage +
                    sqrt_pct_public_coverage_alone + pct_white + pct_married_households +
                    sqrt_birth_rate + log_study_per_cap + log_pct_black + log_pct_asian +
                    log_other_race, 
                    data = model_data_state_mean, 
                    method = "forward",
                    nvmax = 10)
reg_summary_mean = summary(regfit_full_mean)

par(mfrow=c(2, 2))
plot(reg_summary_mean$rss ,xlab="Number of Variables ",ylab="RSS", type="l")

plot(reg_summary_mean$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")

plot(reg_summary_mean$cp,xlab="Number of Variables ",ylab="Cp",type="l")

plot(reg_summary_mean$bic ,xlab="Number of Variables ",ylab="BIC",type="l")

lm_subset = 
```



## Backward subset selection
```{r}
best_backward <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), method = "backward", nvmax=11, ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  


# Select the 'best' model of all subsets for 21-predictor model
round(best_backward(transform_model, nbest = 1), 3)

## plot
regfit_full_mean = regsubsets(target_death_rate ~ log_avg_ann_count + incidence_rate + 
                    inverse_med_income + sqrt_poverty_percent + median_age_male +
                    median_age_female + log_avg_household_size + percent_married +
                    sqrt_pct_no_hs18_24 + pct_hs18_24 + sqrt_pct_bach_deg18_24 + 
                    pct_hs25_over + log_pct_bach_deg25_over + pct_employed16_over +
                    sqrt_pct_unemployed16_over + pct_private_coverage + 
                    pct_private_coverage_alone + pct_emp_priv_coverage + pct_public_coverage +
                    sqrt_pct_public_coverage_alone + pct_white + pct_married_households +
                    sqrt_birth_rate + log_study_per_cap + log_pct_black + log_pct_asian +
                    log_other_race, 
                    data = model_data_state_mean, 
                    method = "backward",
                    nvmax = 10)
reg_summary_mean = summary(regfit_full_mean)

par(mfrow=c(2, 2))
plot(reg_summary_mean$rss ,xlab="Number of Variables ",ylab="RSS", type="l")

plot(reg_summary_mean$adjr2 ,xlab="Number of Variables ",ylab="Adjusted RSq",type="l")

plot(reg_summary_mean$cp,xlab="Number of Variables ",ylab="Cp",type="l")

plot(reg_summary_mean$bic ,xlab="Number of Variables ",ylab="BIC",type="l")
```

forward gives clearest results::
incidence_rate, pct_hs18_24, log_pct_bach_deg25_over, pct_married_households

let's fit the model with the four best

```{r}
best_subset_model = lm(target_death_rate = incidence_rate + pct_hs18_24 + log_pct_bach_deg25_over
                       + pct_married_households, data = )
```




Let's try lasso on the new transformed dataset

```{r}

attach(model_data_state_mean)

library(glmnet)
test.mat = model.matrix(target_death_rate ~ ., data = model_data_state_mean)[, -1]
lasso.mod = glmnet(test.mat, target_death_rate, alpha = 1)
set.seed(1)
cv.out = cv.glmnet(test.mat,target_death_rate,alpha=1)
par(mfrow=c(1, 1))
plot(cv.out)

bestlam = cv.out$lambda.min
bestlam

out = glmnet (test.mat, target_death_rate, alpha=1)
lasso.coef = predict (out, type="coefficients", s= bestlam)
lasso.coef


```


Lasso only removes 4 variables :(

Will not try ridge because we do not want a model with all of the variables, just reduced in coefficient size


```{r}
cv_df =
  crossv_mc(model_data_state_mean, 100) %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(lin_mod = map(train, ~lm(target_death_rate ~ incidence_rate + inverse_med_income + median_age_female + 
                      pct_hs18_24 + pct_hs25_over + pct_bach_deg25_over + sqrt_pct_unemployed16_over + 
                      sqrt_birth_rate + log_pct_black + log_other_race, data = .x)),
         rmse_test_lin_mod = map2_dbl(lin_mod, test, ~rmse(model = .x, data = .y)),
         rmse_train_lin_mod = map2_dbl(lin_mod, train, ~rmse(model = .x, data = .y)))

cv_df %>% 
  dplyr::select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()

mean(cv_df$rmse_train_lin_mod)
mean(cv_df$rmse_test_lin_mod)

```

Quick summary on what's happening. Transformed data is performing best with 4 variables (via backward selection). 
Untransformed data still looks like it's peforming better. 

Haven't tried removing variables due to multicollinearities

Lasso gives poor results; only removes 4 variables; untransformed data removes several more. Untransformed appears 
to be performing better. 

Need to test these models against test data

Four models to try::
1) best subset transformed (v variables)
2) best subset untransformed (5/6 variables)
3) lasso transformed ()
4) lasso untransformed (11 variables)



