---
title: "tian_model_selection"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(broom)
library(HH) 
library(leaps)
library(caret)
library(modelr)
library(faraway)
```

# initial full model
```{r}
data = read_csv("./data/Cancer_Registry.csv") %>% 
  janitor::clean_names()
```

```{r}
model_full_notrans =
  lm(target_death_rate ~ avg_ann_count + incidence_rate + med_income + poverty_percent + study_per_cap + median_age_male + median_age_female + avg_household_size + percent_married + pct_no_hs18_24 + pct_hs18_24 + pct_bach_deg18_24 + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + pct_emp_priv_coverage + pct_black + pct_asian + pct_other_race + pct_married_households + birth_rate, data = data) 
```

Transformation decision mainly based on "margaret_eda.md"
Full model variables exclude variables with high-linearity.
```{r}
data_trans = data %>% 
  mutate(log_avg_ann_count = log(avg_ann_count)) %>% 
  mutate(study_per_cap = recode(study_per_cap, "0" = 0.001)) %>% #fix "-inf" problem
  mutate(log_study_per_cap = log(study_per_cap)) %>% 
  mutate(sqrt_pct_bach_deg18_24 = sqrt(pct_bach_deg18_24)) %>% 
  mutate(log_pct_bach_deg25_over = log(pct_bach_deg25_over)) %>% 
  mutate(sqrt_pct_unemployed16_over = sqrt(pct_unemployed16_over)) %>% 
  mutate(pct_asian = recode(pct_asian, "0" = 0.001)) %>% #fix "-inf" problem
  mutate(log_pct_asian = log(pct_asian)) %>% 
  mutate(pct_other_race = recode(pct_other_race, "0" = 0.001)) %>% #fix "-inf" problem
  mutate(log_pct_other_race = log(pct_other_race)) %>% 
  dplyr::select(target_death_rate, log_avg_ann_count, incidence_rate, med_income, poverty_percent,  log_study_per_cap, median_age_male, median_age_female, avg_household_size, percent_married, pct_no_hs18_24,  pct_hs18_24, sqrt_pct_bach_deg18_24, pct_hs25_over, log_pct_bach_deg25_over, sqrt_pct_unemployed16_over,  pct_emp_priv_coverage, pct_black, log_pct_asian, log_pct_other_race, pct_married_households, birth_rate)

model_full =
  lm(target_death_rate ~ log_avg_ann_count + incidence_rate + med_income + poverty_percent + log_study_per_cap + median_age_male + median_age_female + avg_household_size + percent_married + pct_no_hs18_24 + pct_hs18_24 + sqrt_pct_bach_deg18_24 + pct_hs25_over + log_pct_bach_deg25_over + sqrt_pct_unemployed16_over + pct_emp_priv_coverage + pct_black + log_pct_asian + log_pct_other_race + pct_married_households + birth_rate, data = data_trans) #21 variables
```

# Method 1 stepwise

## without
```{r}
step(model_full_notrans, direction = 'backward')
```

## with transformation
```{r}
step(model_full, direction = 'backward')
```

```{r}
data_15 = data_trans %>% 
   dplyr::select(target_death_rate, log_avg_ann_count, incidence_rate, poverty_percent, median_age_male, percent_married, pct_no_hs18_24,  pct_hs18_24, sqrt_pct_bach_deg18_24, pct_hs25_over, log_pct_bach_deg25_over, sqrt_pct_unemployed16_over, pct_black, log_pct_other_race, pct_married_households, birth_rate)

model_step = lm(formula = target_death_rate ~ log_avg_ann_count + incidence_rate + poverty_percent + median_age_male + percent_married + pct_no_hs18_24 + pct_hs18_24 + sqrt_pct_bach_deg18_24 + pct_hs25_over + log_pct_bach_deg25_over + sqrt_pct_unemployed16_over + pct_black + log_pct_other_race + pct_married_households + birth_rate, data = data_15)
summary(model_step) 
```

delete some variables based on p-value?
```{r}
data_11 = data_15 %>% 
  dplyr::select(-log_avg_ann_count, -pct_no_hs18_24, -sqrt_pct_bach_deg18_24, -pct_black)

model_11 = lm(formula = target_death_rate ~  incidence_rate + poverty_percent + median_age_male + percent_married + pct_hs18_24 + pct_hs25_over + log_pct_bach_deg25_over + sqrt_pct_unemployed16_over + log_pct_other_race + pct_married_households + birth_rate, data = data_11)
summary(model_11) 
```

delete some variables based on high correlation?

```{r}
cor(data_11)
```

cov(pct_married_households, percent_married) = 0.8702605
cov(log_pct_bach_deg25_over, pct_hs25_over) = -0.70434687   
cov(poverty_percent, percent_married) = -0.6428569
cov(poverty_percent, sqrt_pct_unemployed16_over) = 0.630315059

```{r}
data_8 = data_11 %>% 
  dplyr::select(-pct_married_households, -log_pct_bach_deg25_over, -poverty_percent)

model_8 = lm(formula = target_death_rate ~  incidence_rate + median_age_male + percent_married + pct_hs18_24 + pct_hs25_over + sqrt_pct_unemployed16_over + log_pct_other_race + birth_rate, data = data_8)
summary(model_8) 
```

# Method 2 criterion-based (>8?)

```{r criterion}
data_trans = as.data.frame(data_trans)
# Leaps function provides all-subsets analysis
# Printing the 2 best models of each size, using the Cp criterion:
leaps(x = data_trans[,2:9], y = data_trans[,1], nbest = 2, method = "Cp")
# Printing the 2 best models of each size, using the adjusted R^2 criterion:
leaps(x = data_trans[,2:9], y = data_trans[,1], nbest = 2, method = "adjr2")

# Summary of models for each size (one model per size)
rs = summary(regsubsets(target_death_rate ~ ., data = data_8))

par(mfrow = c(1,2))

plot(2:9, rs$cp, xlab = "No of parameters", ylab = "Cp Statistic")
abline(0,1)

plot(2:9, rs$adjr2, xlab = "No of parameters", ylab = "Adj R2")
```

# Method 3 criterion-based (subsets)

```{r}
best <- function(model, ...) {
  subsets <- regsubsets(formula(model), model.frame(model), ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  return(subsets)
}  

# Select the 'best' model of all subsets for 8-predictor model
round(best(model_8, nbest = 1), 8)
```

adjr2 = 0.4465855... not good...

# Check outliers

```{r}
influence.measures(model_8)
```

26   1.27e-04 0.006544   *
31   1.58e-03 0.015489   *
34   3.15e-03 0.004148   *

# Cross-validation

```{r 10_fold}
set.seed(1)
data_train = trainControl(method = "repeatedcv", number = 10, repeats = 10)

# Fit the model
model_caret = train(target_death_rate ~  incidence_rate + median_age_male + percent_married + pct_hs18_24 + pct_hs25_over + sqrt_pct_unemployed16_over + log_pct_other_race + birth_rate, data = data_8, 
                    method = 'lm', na.action = na.pass)
model_caret
```

RMSE:20.85816
Rsquared: 0.4400466

