---
title: "Variable_selection"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
data = read.csv("./Cancer_Registry.csv") %>% 
  janitor::clean_names() 

dim(data)
```

# Check for missing value

```{r}
na_output = vector("list", length = 34)

for (i in 1:34) {
  na_output[[i]] = sum(is.na(data[[i]])) / 3047
}

na_output
```

"na_output" shows the percentages of missing values of each variable.

We should exclude variable 18 (pctsomecol18_24), vairable 25 (pctprivatecoveragealone).

As for variable 22 (pctemployed16_over), 5% is not very high, but it has high collinearity with pctunemployed16_over. So, we decide to delete it anyway.


# Check for collinearity

We can see from the definitions that some variables have similar definations. We can just use a representative one in each group.

For example, "median income" and "median income per capita binned by decile" are similar. We can delete "[9]binnedInc".

For example, "[1]avgAnnCount" & "[2]avgDeathsPerYear" are similar, their correlation  = 0.9394078. We can delete [2]avgDeathsPerYear.

Cov([6]pop_est2015, [1]avg_ann_count) =  0.926893538, Cov([5]pop_est2015, [2]avg_deaths_per_year) = 0.97763406. We can delete [6]pop_est2015.

Similarly, we can delete "[24]pct_private_coverage", "[27]pct_public_coverage", "[28]pct_public_coverage_alone", "[29]pct_white" 

(all based on the correlation coefficients shown below)

```{r}
#pay attention to extrmely high correlation coefficients
cor(data[c(-9, -13, -18, -22, -25)]) #delete [9] and [13] beacause of wrong type.
```


